---
title: "task4_analysis"
output: html_notebook
---
```{r}
library(tidyverse)
library(janitor)
library(tidyr)
library(readxl)
library(lubridate)
library(here)
library(stringr)
```




```{r}
full_candy <- read_csv(here::here("clean_data/full_candy.csv"), show_col_types = FALSE)
```



Question 1
What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
full_candy %>% 
  names()

candies_only <- select(full_candy, -year, -age, -going_out,
                       -country, -gender)
names(candies_only)

sum(!is.na(candies_only))


# [1] 629523

```




Question 2
What was the average age of people who are going out trick or treating?

```{r}
going_out <- subset(full_candy, going_out == "Yes" & age >= 1 & age <= 100)

mean(going_out$age, na.rm = TRUE)

# [1] 35.01952
```


Question 3
What was the average age of people who are not going trick or treating?

```{r}
going_out <- subset(full_candy, going_out == "No" & age >= 1 & age <= 100)

mean(going_out$age, na.rm = TRUE)


# [1] 39.14549
```




Question 4
For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
 # pivoting table to longer

candies_only_long  <- pivot_longer(candies_only,
                                   cols = c(1:97),
                                   names_to = "candy",
                                   values_to = "feelings")

 # grouping by candy

candies_grouped <- candies_only_long %>% 
  group_by(candy)  
candies_grouped <- candies_only_long %>% 
  count(candy, feelings) 


 # selecting for joy

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  slice_max(feelings == "JOY") %>% 
  arrange(desc(n))

 # any_full_sized_candy_bar	JOY	7589	



 # selecting for meh

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  slice_max(feelings == "MEH") %>% 
  arrange(desc(n))

# lollipops	MEH	1570	



# selecting for despair

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  slice_max(feelings == "DESPAIR") %>% 
  arrange(desc(n))

# broken_glow_stick	DESPAIR	7905	


```



Question 5
How many people rated Starburst as despair?
```{r}
candies_grouped %>% 
  select(candy, feelings, n) %>% 
  filter(candy == "starburst", feelings == "DESPAIR")


```



For the next three questions, count despair as -1, joy as +1, and meh as 0.

Question 6
What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r}
candies_only_q6 <- select(full_candy, -year, -age, -going_out,
                       -country)

candies_only_long_q6  <- pivot_longer(candies_only,
                                   cols = c(1:97),
                                   names_to = "candy",
                                   values_to = "feelings"
                                   )
```


