---
title: "task4_analysis"
output: html_notebook
---
```{r}
library(tidyverse)
library(janitor)
library(tidyr)
library(readxl)
library(lubridate)
library(here)
library(stringr)
```




```{r}
full_candy <- read_csv(here::here("clean_data/full_candy.csv"), show_col_types = FALSE)
```



Question 1
What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
full_candy %>% 
  names()

candies_only <- select(full_candy, -year, -age, -going_out,
                       -country, -gender)
names(candies_only)

sum(!is.na(candies_only))


# [1] 629523

```




Question 2
What was the average age of people who are going out trick or treating?

```{r}
going_out <- subset(full_candy, going_out == "Yes" & age >= 1 & age <= 100)

mean(going_out$age, na.rm = TRUE)

# [1] 35.01952
```


Question 3
What was the average age of people who are not going trick or treating?

```{r}
going_out <- subset(full_candy, going_out == "No" & age >= 1 & age <= 100)

mean(going_out$age, na.rm = TRUE)


# [1] 39.14549
```




Question 4
For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
 # pivoting table to longer

candies_only_long  <- pivot_longer(candies_only,
                                   cols = c(1:97),
                                   names_to = "candy",
                                   values_to = "feelings")

 # grouping by candy

candies_grouped <- candies_only_long %>% 
  group_by(candy)  
candies_grouped <- candies_only_long %>% 
  count(candy, feelings) 


 # selecting for joy

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  slice_max(feelings == "JOY") %>% 
  arrange(desc(n))

 # any_full_sized_candy_bar	JOY	7589	



 # selecting for meh

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  slice_max(feelings == "MEH") %>% 
  arrange(desc(n))

# lollipops	MEH	1570	



# selecting for despair

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  slice_max(feelings == "DESPAIR") %>% 
  arrange(desc(n))

# broken_glow_stick	DESPAIR	7905	


```



Question 5
How many people rated Starburst as despair?
```{r}
candies_grouped %>% 
  select(candy, feelings, n) %>% 
  filter(candy == "starburst", feelings == "DESPAIR")


#starburst	DESPAIR	1990	
```



For the next three questions, count despair as -1, joy as +1, and meh as 0.

Question 6
What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r}

# selecting for candies, gender, year and country

candies_only_q6 <- select(full_candy, -age, -going_out)


# changing order so gender column is first

candies_only_q6 <- candies_only_q6 %>% 
  relocate(gender, .before = butterfinger) %>%
   relocate( year, .after = gender) %>% 
   relocate(country, .after = year)
  


# new df with gender, feelings and candy

candies_only_q6_long <- candies_only_q6 %>% 
  pivot_longer(
              cols = c(4:100),
              names_to = "candy",
              values_to = "feelings")


# duplicated feelings column

candies_only_q6_long$rating = candies_only_q6_long$feelings


rating_count_full_candy <- candies_only_q6_long %>% 
  count(candy, rating, year, gender,country)  %>% 
  mutate ( gender = case_when(gender == "I'd rather not say" ~ NA_character_ ,
                              gender == "Other" ~ NA_character_,
                              gender == "NA" ~ NA_character_,
                              TRUE ~ gender))



# converting the rating column to -1. 1, and 0 

rating_count_full_candy_new <- rating_count_full_candy %>% 
  mutate(rating = case_when(rating == "DESPAIR" ~ "-1",
                            rating == "JOY" ~ "1",
                            rating == "MEH" ~ "0",
                            TRUE ~ rating))

 # converting rating column to numeric

rating_count_full_candy_new_numeric <- rating_count_full_candy_new %>% 
  mutate(rating = as.numeric(rating))

# calculating the sum of rating for each candy per gender

popular_candy_by_gender <- rating_count_full_candy_new_numeric %>% 
  filter( rating == -1 | rating == 1 |rating == 0) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candy, gender) %>%
summarise(sum_rating = sum(new_rating)) 
  

# reordering the table descendant to see the highest rated candy

male_popular_candy <- popular_candy_by_gender %>% 
 filter( gender == "Male") %>% 
arrange(desc(sum_rating))

female_popular_candy <- popular_candy_by_gender %>% 
 filter( gender == "Female") %>% 
arrange(desc(sum_rating))

male_popular_candy
female_popular_candy


# most popular
# any_full_sized_candy_bar	Male	1584	
# any_full_sized_candy_bar	Female	875

# second place
# reeses_peanut_butter_cups	Male	1443
# reeses_peanut_butter_cups	Female	768

```



Question 7
What was the most popular candy bar in each year?


```{r}
popular_candy_by_year <- rating_count_full_candy_new_numeric %>% 
  filter( rating == -1 | rating == 1 |rating == 0) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candy, year) %>%
summarise(sum_rating = sum(new_rating)) 

popular_candy_by_year

popular_candy_2015 <- popular_candy_by_year %>% 
 filter(year == "2015") %>% 
arrange(desc(sum_rating))

popular_candy_2016 <- popular_candy_by_year %>% 
 filter(year == "2016") %>% 
arrange(desc(sum_rating))

popular_candy_2017 <- popular_candy_by_year %>% 
 filter(year == "2017") %>% 
arrange(desc(sum_rating))

popular_candy_2015 # 4,603 for any full sized candy bar, 4,375 for reeses peanut candy bar

popular_candy_2016 # 1,037 for any full sized candy bar, 920 for kitkat

popular_candy_2017 # 1,542 for any full sized candy bar, 1,403 for reeses peanut candy bar

# I want to check the answer is correct:
# Overall rating for any full sized candy bar is:

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  filter(candy == "any_full_sized_candy_bar", feelings == "JOY") # 7,589

candies_grouped %>% 
  select(candy, feelings, n) %>% 
  filter(candy == "any_full_sized_candy_bar", feelings == "DESPAIR")# 407 so JOY - DESPAIR = 7,182

# Total of popular candy = 4,603 + 1,037 + 1,542 = 7,182 - YAY!


```



Question 8
What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?


```{r}
# For USA:

popular_candy_by_country <- rating_count_full_candy_new_numeric %>% 
  filter( rating == -1 | rating == 1 | rating == 0 ) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candy, country) %>%
summarise(sum_rating = sum(new_rating)) 

popular_candy_by_country

popular_candy_USA <- popular_candy_by_country %>% 
 filter(country == "USA") %>% 
arrange(desc(sum_rating))

popular_candy_USA # 2,172  for any full sized candy bar, 1,983 for reeses peanut butter cups

popular_candy_UK <- popular_candy_by_country %>% 
 filter(country == "UK") %>% 
arrange(desc(sum_rating))

popular_candy_UK # 33 for any full sized candy bar, lindt truffle, rolos, toblerone

popular_candy_Canada <- popular_candy_by_country %>% 
 filter(country == "Canada") %>% 
arrange(desc(sum_rating))

popular_candy_Canada # 252 for any full sized candy bar, 230 for kitkat


popular_candy_by_country_other <- rating_count_full_candy_new_numeric %>% 
  filter(!is.na(country)) %>%
  filter(country != "USA" & country != "Canada" & country != "UK") %>% 
  filter( rating == -1 | rating == 1 |rating == 0) %>% 
  mutate( new_rating = rating * n ) %>% 
  group_by(candy) %>%
  summarise(sum_rating = sum(new_rating)) %>% 
  arrange(desc(sum_rating))

# most popular candy bar for other countries is any_full_sized_candy_bar	65	


```



